---
title: "simpleRegress"
author: "Israel Valdovinos"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted wieght of a bear compared to chest. adjusted R squared value of 0.9262, and residual standard error value of 33.08.


```{r}
simp2 <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp2)
```

The predicted length of a bear compared to chest. adjusted R squared value of 0.7423, and residual standard error value of 83.83.

```{r}
simp2 <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp2)
```
The predicted length of a bear compared to head width has an adjusted R squared value of 0.76064, and residual standard error value of 76.41.

```{r}
simp4 <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp)
summary.lm(simp4)
```

The predicted length of a bear compared to head length has an adjusted R squared value of 0.69, and residual standard error value of 67.81.

```{r}
simp5 <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp)
summary.lm(simp5)
```
The predicted length of a bear compared to head length has an adjusted R squared value of 0.5526, and residual standard error value of 81.47.